Supplementary figures\nCapturing regeneration in models of forest dynamics

### 2023-04-11 manuscript/supplementaryRW.Rmd
update model references to make it homogeneous across models reports in the suplementary

### Model references

@@

@@

@@

@@

@@

@@

@@

### Model references

Morin X., de Coligny F., Bugmann H., Limousin J.-M., Ourcival J.-M.,
Martin-StPaul N., Simioni G., Cailleret M., Prevosto B., Toïgo M.,
Vennetier M., Cateau E., Guillemot J. 2021. Beyond forest succession: a
gap model to study ecosystem functioning and tree community composition
under climate change. Functional Ecology. 35, 955-975.

Jourdan M., Dreyfus, P., Riond C., Cordonnier T., Cornet B., de Coligny
F., Morin X. 2021. Managing mixed stands can mitigate severe climate
change impacts on French alpine forests. Regional Environmental Change.
21, 78.

Morin X., Damestoy T., Toïgo M., Jactel H., Castagneyrol B. Meredieu C.
2020. Using forest gap models and experimental data to explore long-term
effects of tree diversity on the productivity of mixed planted forests.
Annals of Forest Science. 77
<https://doi.org/10.1007/s13595-020-00954-0>.

@@

### Model references

@@

### Model references

Bugmann, H. (1994). On the ecology of mountainous forests in a changing
climate: A simulation study. Huber, N., Bugmann, H., & Lafond, V.
(2020). Capturing ecological processes in dynamic forest models: Why
there is no silver bullet to cope with complexity. Ecosphere, 11(5).
<https://doi.org/10.1002/ecs2.3109>

@@

Soil data were translated to bucket sizes values based on the
assumpation that the lowest soil quality value of 1 refers to a bucket
size of 10 cm and the highest soil quality of 5 refers to a bucket size
of 35 cm. Specifically we used the following formula to calculate the
bucket size bucket_size = 10 + (soil_quality-1) /4  \* 25.

@@

All simulations were run according to the protocoll. For each site we
simulated on 16 ha (i.e., 100 patches of 0.08 ha). We defined a spin up
phase of 1000 years. After the spin up we simulated 200 years of forest
dynamics.

@@

More detailed information about Sibyla can be found in Fabrika (2005)
and in the link: <http://etools.tuzvo.sk/sibyla/english/model.htm>.

@@

### Model summary

@@

@@

@@

### Model summary

The most complete reference resource for the iLand regeneration module
is: Seidl, R., Spies, T. a. T. A. T. a., Rammer, W., Steel, E. A. A.,
Pabst, R. J. R. J. R. J., & Olsen, K. (2012). Multi-scale Drivers of
Spatial Variation in Old-Growth Forest Carbon Density Disentangled with
Lidar and an Individual-Based Landscape Model. Ecosystems, 15(8),
1321-1335. <https://doi.org/10.1007/s10021-012-9587-2>

@@

@@

The most complete reference resources for the LANDIS-II model and
regeneration module used here are:

@@

@@

• Bugmann, H. 1994. On the ecology of mountainous forests in a changing
climate: A simulation study. Dissertation. Swiss Federal Institute of
Technology, Zurich. • Fischlin, A., H. Bugmann, and D. Gyalistras. 1995.
Sensitivity of a forest ecosystem model to climate parametrization
schemes. Env. Poll. 87:267-282. • Lischke, H., and T. J. Loffler. 2006.
Intra-specific density dependence is required to maintain species
diversity in spatio-temporal forest simulations with reproduction.
Ecological Modelling 198:341-361. • Lischke, H., T. J. Loffler, and A.
Fischlin. 1998. Aggregation of individual trees and patches in forest
succession models: Capturing variability with height structured, random,
spatial distributions. Theor Popul Biol 54:213-226. • Lischke, H., N. E.
Zimmermann, J. Bolliger, S. Rickebusch, and T. J. Loffler. 2006.
TreeMig: A forest-landscape model for simulating spatio-temporal
patterns from stand to landscape scale. Ecological Modelling
199:409-420. • Thornthwaite, C. W., and J. R. Mather. 1957. The water
balance. Publications in Climatology VIII:1-69.

@@

We used the process-based mechanistic model LPJ-GUESS to simulate
vegetation dynamics considering tree species, age cohorts, gap dynamics
and biogeochemical cycles. Parameters of the model reflect an updated
version of the most common European tree species as well as typical
shrub PFTs as described by Hickler et al. (2012). A detailed description
of LPJ-GUESS is available in Smith et al. 2001 and Smith et al. 2014.
Source code is available on demand (see
<https://web.nateko.lu.se/lpj-guess/>).

@@

All sites We used were simulated withdefault values for a medium
textured soil type (sand: 0.35, clay: 0.15, silt: 0.5) in LPJ-GUESS.

@@

>>>>

Mahnken M., Gutsch M., Reyer C.P.O., Lasch-Born P.

The most complete reference resource for the 4C (v2.2) model and
regeneration module is:

Lasch-Bonn et al. 2020. Description and evaluation of the process-based
forest model 4C v2.2 at four European forest sites. DOI:
[https://doi.org/10.5194/gmd-13-5311-2020)](https://doi.org/10.5194/gmd-13-5311-2020))

@@

We used the provided daily ERA5-CHELSA climate data set and complemented
the CHELSA time series data regarding relative humidity with the model
after Eccel et al. (2015) based on minimum and mean daily temperature to
accomodate data needs of 4C. The missing data for January 2nd and 3rd
2013 were filled with data from January 1st and 4th 2013 respectively.
For construction of longer climate time series we looped through the
provided time series from 1981 to 2018 up to the maximum simulation
length of 2500 years. We amended situations where leap year
inconsistencies arised by removing February 29 or duplicating data from
February 28. Otherwise, the unaltered daily data were used to run 4C on
a daily basis. We applied a constant atmopsheric CO2 concentration of
380 ppm.

@@

We used parametrized soils readily available for 4C from the EU soil
data base. The link from parametrized soil to soil quality was done by
extracting the plant available water storage capacity as the model
bucket size of the rooted soil horizons. We excluded bucket size values
of lower than 10 cm and larger than 35 cm. The bucket size was
reprojected to the range from 0 to 1. In addition we tested the
sensitivity of dominant tree height at age 100 years for monospecific
stands of all five species descibed above to the available soils using
model simulations with 4C. The dominant tree height at age 100 years was
then also reprojected to the range from 0 to 1 in order to derive a
model specific soil quality indicator for all paramterized soils. Then
we multiplied this relative model specific soil quality with the
relative soil quality derived from the bucket size to the range from 1
to 5. The match between paramterized soil and plot was then done by
finding the pair of soil quality with the minimum difference between
simuation protocol prescribed soil quality and and the soil quality of
the paramterized soils as described above.

@@

Topographical information was not considered in the simulations.

@@

We initialized stands with five out of the eleven proposed
species/genera: Fagus sylvatica (fasy), Picea abies (piab), Betula spp.
(betu; 4C parameters for Betula pendula), Quercus spp. (quer; 4C
parameters for Quercus robur) and Pinus sylvestris (pisy) since only
these are fully parameterized for an application as proposed in the
simulation protocol. Each plot was initialized with all five tree
species with similar distributions of height classes and number of
individuals: 2000 individuals/ha with height around 0.25 m, 1000
individuals/ha with height around 0.75 m and 500 individuals/ha with
height around 1.75 m. In total 17500 individuals/ha.
Carpinus betulus,
Tilia cordata, Acer pseudoplatanus, Fraxinus excelsior, Alnus glutinosa
and Abies alba are not fully parameterized in 4C.

@@

#### Dispersal and seed input

There is no explicitly modelled seed dispersal in a stand and between
stands. We applied a constant sapling establishment rate every five
years of 2000 individuals/ha for all five species; 10000 individuals/ha
in total every five years. The constant sapling establishment provides
constant regeneration potential if favorable conditions for sapling
growth (especially light and water regime) are present. This approach
enables regeneration of other species than only the one present in the
stand to establish if competition allows. The resulting recruitment
patterns are an outcome of the dynamical processes from the sapling
stage onward. Seed production and germination were not explicitly
simulated.

#### Simulation length and equilibrium criteria

To reach equilibrium and obtain dynamics in the equilibrium period, we
simulated a maximum of 2500 years. 4C is usually applied for much
shorter simulation periods and different technical as well as
model-theory related issues arise when simulating longer than 100-150
years. Therefore 2500 years can be seen as the upper limit of continuous
simulation in the defined simulation setup. The first 500 years were not
sampled for recruitment as we assume the model needs this time to reach
equilibrium. In this long simulation period 4C partially projects
unrealistically large tree heights resulting in premature termination of
the simulation (36 plots) so that not all plots were simulated for the
full 2500 years. We started sampling the recruits after the first 500
simulation years from non-overlapping 10 year periods to obtain 200
samples. For shorter simulation periods we produced 200 samples from
overlapping 10 year periods.
In addition, we excluded those plots in which 4C simulated
unrealistically high basal area values of over 150 m² (19 plots). We
used this large threshold to give the model maximum flexibility to
explore the solution space in these long simulation runs although we
acknowledge that this is an unrealistic overestimation that is probably
originating from misrepresented density dependent mortality processes in
long simulation runs and misrepresented single tree dimensions due to
wrong assumptions of tree geometry and biomass share of such old-aged
individuals. Furthermore those plots with less than 1000 simulation
years (16 plots) before termination were excluded because the sample
space after the first 500 years was too small. Additionally we included
another subset criterion removing all samples that have a ba \> 100 to
exclude unrealistically high samples without the need to exclude the
full site. This resulted in discarding 0.8 % of the sample

@@

We provide 200 samples each for 165 of the 200 plots (see reason for
excluded plots in "Simulation length and equilibrium criteria"). These
stem from 10 year periods that overlap partially (for \< 2500 simulation
years) or do not overlap (for 2500 simulation years). The number of
recruits (r.trees) is the sum of individuals that cross the threshold
diameter (7/10 cm) in the 10 year sample period per hectare that have
not already died until the sample year (Fig. 3). The recruit's basal
area (r.ba) is the sum of the basal area per hectare of the trees that
cross the diameter threshold that have not already died until the sample
year. The total stand basal area (ba) is the basal area per hectare of
all trees in the stand \> 0 cm diameter at breast height (no lower
boundary) at the sample time per species. We provide the data for both
diameter thresholds, 7 and 10 cm. In over 90 % of the decades no
recruitment takes place.

[4C
(v2.2)]([https://doi.org/10.5194/gmd-13-5311-2020)](https://doi.org/10.5194/gmd-13-5311-2020))
is usually not used for simulations of forest equillibrium dynamics,
potential natural vegetation assessment and long simulation periods but
rather shorter-term simulations (up to 100-150 years). In addition, only
five species out of the pool of eleven potentially relevant species here
are parameterized in 4C so that projections of the potential natural
vegetation done wih 4C need to be considered in the light of model
uncertainty.
For example, the spatial homogeneity in the model leads to
cyclic recruitment patterns (see Fig. 4). When the adult stands dies of,
young individuals are recruited into the larger diameter classes because
more light reaches lower canopy layer since larger (adult) trees do not
anymore prevent the light reaching lower canopy layers. This results in
increased growth of the regeneration layer during those times. In most
stands Fagus sylvatica dominates the stand after some time but
occasionally individuals of other species, like Picea abies or Quercus
spp. cross the recruitment diameter threshold.

@@

The most complete reference resource for the ForCEEPS & ForCEEPS(f)
model and regeneration module are:

-   Morin X., de Coligny F., Bugmann H., Limousin J.-M., Ourcival J.-M.,
    Martin-StPaul N., Simioni G., Cailleret M., Prevosto B., Toïgo M.,
    Vennetier M., Cateau E., Guillemot J. 2021. Beyond forest
    succession: a gap model to study ecosystem functioning and tree
    community composition under climate change. Functional Ecology. 35,
    955-975.

-   Jourdan M., Dreyfus, P., Riond C., Cordonnier T., Cornet B., de
    Coligny F., Morin X. 2021. Managing mixed stands can mitigate severe
    climate change impacts on French alpine forests. Regional
    Environmental Change. 21, 78.

-   Morin X., Damestoy T., Toïgo M., Jactel H., Castagneyrol B.
    Meredieu C. 2020. Using forest gap models and experimental data to
    explore long-term effects of tree diversity on the productivity of
    mixed planted forests. Annals of Forest Science. 77
    <https://doi.org/10.1007/s13595-020-00954-0>.

@@

@@

The most complete reference resource for the ForClim 1 & ForClim 11
model and regeneration module is:

-   Bugmann, H. (1994). On the ecology of mountainous forests in a
    changing climate: A simulation study. Huber, N., Bugmann, H., &
    Lafond, V. (2020). Capturing ecological processes in dynamic forest
    models: Why there is no silver bullet to cope with complexity.
    Ecosphere, 11(5). <https://doi.org/10.1002/ecs2.3109>

@@

Soil data were translated to bucket sizes values based on the assumption
that the lowest soil quality value of 1 refers to a bucket size of 10 cm
and the highest soil quality of 5 refers to a bucket size of 35 cm.
Specifically we used the following formula to calculate the bucket size
bucket_size = 10 + (soil_quality-1) /4  \* 25.

@@

All simulations were run according to the protocol. For each site we
simulated on 16 ha (i.e., 100 patches of 0.08 ha). We defined a spin up
phase of 1000 years. After the spin up we simulated 200 years of forest
dynamics.

@@

The most complete reference resource for the SIBYLA model and
regeneration module is:

-    Fabrika (2005) and in the link:
    <http://etools.tuzvo.sk/sibyla/english/model.htm>.

@@

@@

The most complete reference resource for the xComp model and
regeneration module is:

-

@@

The most complete reference resource for the PICUS model and
regeneration module is:

-

@@

The most complete reference resource for the iLand regeneration module
is:

-   Seidl, R., Spies, T. a. T. A. T. a., Rammer, W., Steel, E. A. A.,
    Pabst, R. J. R. J. R. J., & Olsen, K. (2012). Multi-scale Drivers of
    Spatial Variation in Old-Growth Forest Carbon Density Disentangled
    with Lidar and an Individual-Based Landscape Model. Ecosystems,
    15(8), 1321-1335. <https://doi.org/10.1007/s10021-012-9587-2>

@@

The most complete reference resource for the LandClim model and
regeneration module is:

-

@@

The most complete reference resources for the LANDIS-II model and
regeneration module are:

@@

The most complete reference resource for the TreeMig model and
regeneration module are:

-   Lischke, H., T. J. Loffler, and A. Fischlin. 1998. Aggregation of
    individual trees and patches in forest succession models: Capturing
    variability with height structured, random, spatial distributions.
    Theor Popul Biol 54:213-226.

-   Lischke, H., N. E. Zimmermann, J. Bolliger, S. Rickebusch, and T. J.
    Loffler. 2006. TreeMig: A forest-landscape model for simulating
    spatio-temporal patterns from stand to landscape scale. Ecological
    Modelling 199:409-420.

-   Lischke, H., and T. J. Loffler. 2006. Intra-specific density
    dependence is required to maintain species diversity in
    spatio-temporal forest simulations with reproduction. Ecological
    Modelling 198:341-361.

-   The online resources available at treemig.wsl.ch

@@

Bugmann, H. 1994. On the ecology of mountainous forests in a changing
climate: A simulation study. Dissertation. Swiss Federal Institute of
Technology, Zurich.

Fischlin, A., H. Bugmann, and D. Gyalistras. 1995. Sensitivity of a
forest ecosystem model to climate parametrization schemes. Env. Poll.
87:267-282.

Lischke, H., and T. J. Loffler. 2006. Intra-specific density dependence
is required to maintain species diversity in spatio-temporal forest
simulations with reproduction. Ecological Modelling 198:341-361.

Lischke, H., T. J. Loffler, and A. Fischlin. 1998. Aggregation of
individual trees and patches in forest succession models: Capturing
variability with height structured, random, spatial distributions. Theor
Popul Biol 54:213-226.

Lischke, H., N. E. Zimmermann, J. Bolliger, S. Rickebusch, and T. J.
Loffler. 2006. TreeMig: A forest-landscape model for simulating
spatio-temporal patterns from stand to landscape scale. Ecological
Modelling 199:409-420.

Thornthwaite, C. W., and J. R. Mather. 1957. The water balance.
Publications in Climatology VIII:1-69.

@@

We used the process-based mechanistic model LPJ-GUESS to simulate
vegetation dynamics considering tree species, age cohorts, gap dynamics
and biogeochemical cycles. Parameters of the model reflect an updated
version of the most common European tree species as well as typical
shrub PFTs as described by Hickler et al. (2012).

The most complete reference resource for the LPJ-GUESS model and
regeneration module is:

-   Smith, B. (2001). LPJ-GUESS-an ecosystem modelling framework.
    Department of Physical Geography and Ecosystems Analysis, INES,
    Sölvegatan, 12, 22362.

-   Smith, B., Wårlind, D., Arneth, A., Hickler, T., Leadley, P.,
    Siltberg, J., & Zaehle, S. (2014). Implications of incorporating N
    cycling and N limitations on primary production in an
    individual-based dynamic vegetation model. Biogeosciences, 11(7),
    2027-2054.

-   Source code is available on demand (see
    <https://web.nateko.lu.se/lpj-guess/>).

@@

All sites We used were simulated with default values for a medium
textured soil type (sand: 0.35, clay: 0.15, silt: 0.5) in LPJ-GUESS.

@@

The most complete reference resource for the aDGVM2 model and
regeneration module is:

-   Scheiter S, Langan L, Higgins SI (2013) Next generation dynamic
    global vegetation models: learning from community ecology. NEW
    PHYTOLOGIST, 198, 957-969.
